---
title: "Coronavirus (COVID-19) Visualization"
author: "dillonchewwx"
date: "04/04/2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

# Overview

During this pandemic, most of the reports which we get are often through the media. However, these reports tend to be inaccurate and inconsistent as they tend to cherry-pick the statistics and graphs to generate attention grabbing headlines. As such, to get an accurate and reliable picture of the current impact of COVID-19, we should pull the data directly from a reliable source and analyze it. In this report, we will obtain data directly from [Our World in Data](https://ourworldindata.org/coronavirus-source-data) which relies on data from [Johns Hopkins University](https://coronavirus.jhu.edu/map.html) and visualize them using the `plotly` and `leaflet` packages in R. 

## Load Packages
```{r Load Packages, message=FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(leaflet)
library(lubridate)
library(plotly)
library(rgdal)
```

## Load and Explore Data
```{r Load Data, warning=FALSE, message=FALSE}
# Load Data
data<-read_csv("https://covid.ourworldindata.org/data/owid-covid-data.csv")
glimpse(data)
```
To visualize the most updated data, we shall use a Choropleth Map to show the number of events (e.g. Cases, Deaths etc.) for each country. 
```{r Plotly Choropleth, warning=FALSE}
data_today<-data %>%
  filter(date==today()-2) %>%
  filter(!str_detect(iso_code, "OWID"))

plot_ly(type="choropleth",
        locations=data_today$iso_code,
        z=data_today$total_cases,
        text=data_today$location,
        color=data_today$total_cases, 
        colors="Reds") %>%
  colorbar(title="Number of Cases", 
           y=0.75) %>%
  layout(title=list(text=paste0("Number of COVID-19 cases as of ", today()-2), y=0.9))
```
Using the `leaflet` package, we can visualize the data in a similar fashion. We start by first downloading the shapefile and loading the data.
```{r Download Shapefile, warning=FALSE}
if(!file.exists("Data/TM_WORLD_BORDERS_SIMPL-0.3.shp")){
  download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip", 
              destfile="Data/world_shape_file.zip")
}

world<-readOGR( 
  dsn=paste0(getwd(), "/Data"), 
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE
)

# Replace world_spdf@data with our data
world@data<-world@data %>% 
  select(ISO3) %>% 
  rename(iso_code=ISO3) %>% 
  left_join(data_today, by="iso_code")
```
We shall now plot the choropleth.
```{r Leaflet Choropleth, warning=FALSE}
mybins<-c(0, 10000, 50000, 100000, 500000, 10000000, 20000000, 40000000)
mypal<-colorBin( palette="YlOrRd", domain=world@data$total_cases, na.color="transparent", bins=mybins)

mytext<-paste(
    "Country: ", world@data$location,"<br/>", 
    "Total Cases: ", world@data$total_cases, "<br/>", 
    "Total Deaths: ", world@data$total_deaths, "<br/>",
    "Population: ", round(world@data$population, 2), 
    sep="") %>%
  lapply(htmltools::HTML)

leaflet(world) %>% 
  addTiles() %>%
  addPolygons(fillColor=~mypal(total_cases), 
              stroke=TRUE, 
              fillOpacity=0.9,
              color="white", 
              smoothFactor=0.5,
              weight=0.3, 
              label=mytext, 
              labelOptions=labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>% 
  addLegend(pal=mypal, 
            values=~total_cases, 
            opacity=0.9, 
            title=paste0("Number of Cases as of ", today()-2), 
            position="bottomleft")
```
