---
title: "Coronavirus (COVID-19) Visualization"
author: "dillonchewwx"
date: "04/04/2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

# Overview

During this pandemic, most of the reports which we get are often through the media. However, these reports tend to be inaccurate and inconsistent as they tend to cherry-pick the statistics and graphs to generate attention grabbing headlines. As such, to get an accurate and reliable picture of the current impact of COVID-19, we should pull the data directly from a reliable source and analyze it. In this report, we will obtain data directly from [Our World in Data](https://ourworldindata.org/coronavirus-source-data) which relies on data from [Johns Hopkins University](https://coronavirus.jhu.edu/map.html) and visualize them using the `plotly` and `leaflet` packages in R. 

## Load Packages
```{r Load Packages, message=FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(leaflet)
library(lubridate)
library(plotly)
```

## Load and Explore Data
```{r Load Data, warning=FALSE, message=FALSE}
# Load Data
data<-read_csv("https://covid.ourworldindata.org/data/owid-covid-data.csv")
glimpse(data)
```
To visualize the data, we shall use a Choropleth Map to show the number of events (e.g. Cases/Deaths) for each country. 
```{r Tree Plot, fig.height=10, fig.width=10}
plot_ly(data, 
        type="choropleth", 
        locations=data$iso_code, 
        z=data$total_cases, 
        text=data$location, 
        colorscale="Blues")
```
